---
title: "AOV"
output: html_notebook
---

AOV Exercises

```{r}
y = c(2.3, 1.2, 2.9, 2, 1.5, 5, 5.8, 6.3, 6.5, 7, 3.4, 4.1, 3.9, 3.1, 4.5)
```
```{r}
ylab = c(rep("A",5), rep("B", 5), rep("C", 5))
```

```{r}
d2 = data.frame(y, factor(ylab))
d2
```

```{r}
colnames(d2) = c("y", "group")
```

```{r}
y.aov = aov(y~group, data = d2)
summary(y.aov)
```
The degrees of freedom of group should be k-1 = 2.

Since the p value is less than 0.05, we reject the null hypothesis to conclude that at least one mean is different from the rest. 
In anova, we don't test the variance itself. We use the variance to test the means.

H0: mu_A = mu_B = mu_C
H1: at least one mu is different
```{r}
summary.lm(y.aov)
```

We are checking groupA. 
In the first one, we are testing mu_A - mu_B = 0. We reject H0 and conclude that mu_A is not equal to mu_B. 
In the second one, we are testing mu_A - mu_C = 0. We reject H0 and conclude that mu_A is not equal to mu_C. 

```{r}
d2$group = relevel(d2$group, ref=2) #rearrange factors
d2.aov2 = aov(y~group, data=d2)
summary(d2)
summary.lm(d2.aov2) # testing contrasts
```



