---
title: "R Notebook"
output: html_notebook
---

Bataller 2018
```{r}
y = c(21.1, 23.6, 19.3,33,28.6,35,32,26.8,38.6,21.7,15.7,20.6,41.8,36.7,28.4,23.6,31.8,20.7,22.8,32.8)
x0 = c(rep(1,20))
x1 = c(8,5,2,15,11,14,9,7,22,3,1,5,23,17,12,14,8,4,2,8)
x2 = c(35,43,51,60,73,80,76,54,55,90,30,44,84,76,68,25,90,62,80,72)
X = cbind(x0,x1,x2)
```


```{r}
XtXi = solve(t(X)%*%X)
Xty = t(X)%*%y
beta = XtXi%*%Xty
XtXi
Xty
beta
```
```{r}
H = diag(20) - X%*%XtXi%*%t(X)
c = matrix(c(0,1,0,0,0,1), nrow = 2, ncol = 3, byrow = TRUE)
c
```

```{r}
Fnum = t(c%*%beta)%*%solve(c%*%XtXi%*%t(c))%*%(c%*%beta)
Fdenom = t(y)%*%H%*%y
R2 = Fnum/(Fnum+Fdenom)
R2
```

Ftest 
Test the null hypothesis that all the coefficients are zero.

```{r}
c = matrix(c(0,1,0,0,0,1), nrow = 2, ncol =3, byrow = TRUE)
Fnum = t(c%*%beta)%*%solve(c%*%XtXi%*%t(c))%*%(c%*%beta)
Fdenom = t(y)%*%H%*%y
df1 =2
df2 = 17
F.all = (Fnum/df1)/(Fdenom/df2)
F.all
pval = 1-pf(F.all,df1,df2)
pval 
```
Since p < 0.05, we reject the null hypothesis and conclude that at least one variable is nonzero. 

test significance of the intercept
```{r}
library(Matrix)
c = t(c(1,0,0))
Fnum = t(c%*%beta)%*%solve(c%*%XtXi%*%t(c))%*%(c%*%beta)
Fdenom = t(y)%*%H%*%y
df1 = 1
df2 = 17
F.all = (Fnum/df1)/(Fdenom/df2)
F.all
pval = 1-pf(F.all,df1,df2)
pval 
```
Since the p value is less than 0.05, the intercept is significant. 

Check the significance of the variables

```{r}
library(Matrix)
c = t(c(0,1,0))
Fnum = t(c%*%beta)%*%solve(c%*%XtXi%*%t(c))%*%(c%*%beta)
Fdenom = t(y)%*%H%*%y
df1 = 1
df2 = 17
F.all = (Fnum/df1)/(Fdenom/df2)
F.all
pval = 1-pf(F.all,df1,df2)
pval 
```
```{r}
library(Matrix)
c = t(c(0,0,1))
Fnum = t(c%*%beta)%*%solve(c%*%XtXi%*%t(c))%*%(c%*%beta)
Fdenom = t(y)%*%H%*%y
df1 = 1
df2 = 17
F.all = (Fnum/df1)/(Fdenom/df2)
F.all
pval = 1-pf(F.all,df1,df2)
pval 
```
Verify using R function

```{r}
data1 = data.frame(y,x1,x2)
model = lm(y~., data = data1)
summary(model)
```

Since all the variables have a pvalue less than 0.05, all the variables are significant. 

F. Determine a 95% C.I. for the coefficients including the intercept

```{r}
data = data.frame(y,x1,x2)
```

```{r}
summary(lm(y~., data = data))
```

F.
Since the p value of all three variables are below 0.05, we conclude that all three variables are significant.

Confidence Intervals are:

intercept: (9.2535, 10.9779)
B_1: (0.8529, 0.9324)
B_2: (0.1334, 0.1593)

G. Find point estimate if teacher has 20 years of experience and principal's rating is 95

Point Estimate: 41.8748
