---
title: "pset no.5 chan shio"
output: html_notebook
---



```{r}
yA = c(2.3,1.2,2.9,2.0,1.5)
yB = c(5.0,5.8,6.3,6.5,7.0)
yC = c(3.4,4.1,3.9,3.1,4.5)
yD = c(4.6,4.8,5.0,5.2,4.9)
y = c(yA,yB,yC,yD)
n = length(yA)
k = 4
xA = c(rep(1,n), rep(0,n), rep(0,n), rep(0,n))
xB = c(rep(0,n), rep(1,n), rep(0,n), rep(0,n))
xC = c(rep(0,n), rep(0,n), rep(1,n), rep(0,n))
xD = c(rep(0,n), rep(0,n), rep(0,n), rep(1,n))
X = cbind(xA,xB, xC, xD)
```

```{r}
X
```
```{r}
df1 = k-1
df2 = k*(n-1)
XtXi = matrix(rep(0, k*k), nrow = k)
for(i in 1:k){
  XtXi[i,i] = 1/n
}
XtXi
```

```{r}
beta = XtXi %*% t(X) %*%y
beta
```
```{r}
SSE = t(y)%*%y - t(beta)%*%t(X)%*%y
s2 = SSE/df2
```
Hypothesis testing for Beta: At least two of the means are unequal.
mu_1 = mu_2 = mu_3 = mu_4
```{r}
C = matrix(c(1,-1,0,0,1,0,-1,0,1,0,0,-1), nrow = 3, ncol = 4, byrow = TRUE)
C
```
```{r}
Fnum = t(C%*%beta)%*%solve(C%*%XtXi%*%t(C))%*%(C%*%beta)
Fnum
Fdenom = SSE
F.all = (Fnum/df1)/(Fdenom/df2)
F.all
#F test
p = 1 - pf(F.1,df1,df2)
p
```

Since p < 0.05, we reject the null hypothesis and conclude that at least one mean is different from the other means.

E. Test mu_A = mu_B

```{r}
c = t(c(1,-1,0,0))
df1.1 = 1
Fnum = t(c%*%beta)%*%solve(c%*%XtXi%*%t(c))%*%(c%*%beta)
Fdenom = SSE
F.1=(Fnum/1)/(Fdenom/df2)
F.1
pval = 1-pf(F.1,df1.1,df2)
pval
```

Since p < 0.05, we reject the null hypothesis. We conclude that that mu_A is not equal to mu_B. 

Testing another contrast mu_B = mu_C
```{r}
c = t(c(0,1,-1,0))
df1.1 = 1
Fnum = t(c%*%beta)%*%solve(c%*%XtXi%*%t(c))%*%(c%*%beta)
Fdenom = SSE
F.1=(Fnum/1)/(Fdenom/df2)
F.1
pval = 1-pf(F.1,df1.1,df2)
pval
```


To verify:
```{r}
ylab = c(rep("A",5), rep("B",5), rep("C", 5), rep("D", 5))
d2 = data.frame(y, factor(ylab))
colnames(d2) = c("y", "group")
y.aov = aov(y~group, data = d2)
summary(y.aov)
```
```{r}
summary.lm(y.aov)
```
```{r}
d2$group = relevel(d2$group, ref = 2)
d2.aov2 = aov(y~group, data = d2)
summary(d2.aov2)
summary.lm(d2.aov2)
```

Manual Regression

```{r}
data = read.csv("PS3.csv")
data = data[-1]
data
```

```{r}
ts = ts(data["y"])
plot(ts)
```

```{r}
library(TSA)
library(tseries)
Box.test(ts, type = "Ljung")
adf.test(ts)
```
The time series data is autocorrelated and stationary since the p values under both tests are below alpha = 0.05.


```{r}
x0 = c(rep(1,147))
X = cbind(x0, data[,-1])
X = as.matrix(X)
X
```
```{r}
y = data[1]
y = as.matrix(y)
XtX = t(X) %*% X
XtXi = solve(XtX)
XtXi
```


```{r}
beta = XtXi%*%t(X)%*%y
beta
```
```{r}
H = diag(147)-X%*%XtXi%*%t(X)
e = H%*%y
e
```

