---
title: "Data Analysis Project"
author: "Brynx Junil Alegarbes, Neil Bryant Que, Jeremy Marcus Tan"
date: "2024-07-03"
output: pdf_document
---

```{r}
library(TSA)
library(tseries)
library(forecast)
library(fGarch)
```

```{r}
data=read.csv("2022_sto_nino.csv")
X=ts(data)

```

```{r}
plot(X)
```

```{r}
adf.test(X)
```
#Since the p-value is less than 0.05, we reject the null hypothesis and by the ADF test, the process is stationary.

```{r}
Box.test(X, type='Ljung')
```
#Since the p-value is less than 0.05, we reject the null hypothesis and by the Ljung-Box test, the process has serial correlation.

```{r}
acf(X, lag=20, main='ACF Correlogram')
```
#The ACF does not seem to cut off so no MA models can be tested.

```{r}
pacf(X, lag=20, main='PACF Correlogram')
```
#The possible lag orders where the PACF cuts off are 4 and 10. However, we still wanted to check the lag order 3 for parsimony. Moreover, models with lag orders past 10 will not be checked because they would be too complex and too many parameters to estimate. Hence, we will test the models AR(3), AR(4), and AR(10).

```{r}
Arima(X, order=c(3,0,0))
Arima(X, order=c(4,0,0))
Arima(X, order=c(10,0,0))
```
#By the three information criteria and the log likelihood, AR(4) is better than AR(3). Despite improvement in the log likelihood, AIC, and AICc for the AR(10), they are minimal compared to the additional parameters that will be estimated by the model. To maintain parsimony, AR(4) will be chosen as the best model.

```{r}
model=Arima(X, order=c(4,0,0))
```

```{r}
tsdiag(model)
```
#Through the ACF and p value plot, the residuals do not seem to exhibit autocorrelation.

```{r}
Box.test(model$residuals, type='Ljung')
```
#Through the Ljung-Box test, it is further verified that the residuals are not serially correlated.

```{r}
Y2=residuals(model)**2
```

```{r}
Box.test(Y2, type='Ljung')
```
#However, by the Ljung-Box test, the squared residuals are not serially correlated. This means that the residuals are autoregressive conditionally heteroscedastic (ARCH).

```{r}
pacf(Y2, lag=100) #ARCH(1)
```
#Through the PACF of the squared residuals, the only ARCH model that is parsimonious and can be tested is ARCH(1).

#Various hydrologic models used time series models to evaluate volatility and eliminate the ARCH effect. In modeling rainfall–runoff relationships, Modarres and Ouarda (2013) used GARCH(1,1) and ARCH(2) models. Moreover, in predicting daily streamflow, Wang, et al. (2023) used the GARCH(1,1) model. Lastly, in modeling streamflow Fathian, et al. (2019) found the ARCH(5) and ARCH(7) models to be the most fitted to their residual series.

#Hence, combining the models above, we will evaluate the volatility models and perform joint estimation with our best ARIMA model.

```{r}
arch1 = garchFit(~arma(4,0)+garch(1,0),data=X)
arch2 = garchFit(~arma(4,0)+garch(2,0),data=X)
arch5 = garchFit(~arma(4,0)+garch(5,0),data=X)
arch7 = garchFit(~arma(4,0)+garch(7,0),data=X)
garch11 = garchFit(~arma(4,0)+garch(1,1),data=X)
```
```{r}
arch1 = Arima(Y2, order=c(1,0,0))
arch2 = Arima(Y2, order=c(2,0,0))
arch5 = Arima(Y2, order=c(5,0,0))
arch7 = Arima(Y2, order=c(7,0,0))
garch11 = Arima(Y2, order=c(1,0,1))
```

```{r}
arch1
arch2
arch5
arch7
garch11
```

#Best model is AR(4) - ARCH(1)

```{r}
#Forecastin
```

