---
title: "BDO Daily"
output: html_notebook
---

Loading Packages
```{r}
library("TSA")
library("tseries")
library("lmtest")
library("tidyverse")
library("forecast")
library("quantmod")
library("ggplot2")
```

Source of Data:
Yahoo Finance 
Wall Street Journal
Daily closing prices from May 15, 2018 to May 11, 2023

Preprocessing the Data

Sort data from oldest to newest
```{r}
# Read the CSV data
BDO = read.csv("BDO_daily.csv")


#Sort data from oldest to newest
BDO = BDO[rev(rownames(BDO)), ]
BDO = tail(BDO, 500)

training_BDO = head(BDO, 470)
training_BDO 


test_BDO = BDO
```



Create time series data
```{r}
# training data
BDO = training_BDO["Close"] 
BDO_TS = ts(BDO)
BDO_TS
plot(BDO_TS)
```


Check for stationarity
```{r}
adf.test(BDO_TS)
```


Pre-process the price data. 
```{r}
BDO_TS_diff = diff(BDO_TS) # differencing
adf.test(BDO_TS_diff)
```
Since p<0.05, the price data is stationary. 

Check for autocorrelation
```{r}
Box.test(BDO_TS_diff, type = "Ljung")
```
Since p<0.05, we conclude that the price data is autocorrelated.

Modeling
```{r}
acf(BDO_TS_diff, main = "ACF") # Check MA(3)
acf(BDO_TS_diff, type="partial", main = "PACF") # Check AR(3), AR(4)
```

We check the following models:

Lower AIC indicates the best fitting model.

```{r}
arima(BDO_TS_diff, order=c(3,0,0)) 
arima(BDO_TS_diff, order=c(4,0,0)) 
arima(BDO_TS_diff, order=c(0,0,3)) 
arima(BDO_TS_diff, order=c(3,0,3)) 
arima(BDO_TS_diff, order=c(4,0,3)) 

```
According to AIC,
our best fitting model is ARMA(4,0)

Define the models
```{r}
arma_40 = Arima(BDO_TS_diff, order=c(4,0,0))
arma_40
```

Check the residuals. We want to see if they are autocorrelated.
```{r}
checkresiduals(arma_40, test = "LB")
```
Since the p value > 0.05, we conclude that the residuals are not autocorrelated. (which is what we want)
Looking at the correlograms for the ACF and PACF, there are only a few lags which exceed the confidence threshold so the correlation between the residuals is not significant. The graph of the residuals shows that the residuals have a mean of around 0. 




```{r}
# forecast the actual prices
forecast_40 = forecast(arma_40, level = c(95), h = 30)
plot(forecast_40)
```

```{r}
forecast_40
```

```{r}
accuracy(forecast_40)
```
```{r}
best = Arima(BDO_TS, order = c(4,1,0))
forec = forecast(best, level = c(95), h = 30 )
forec
plot(forec)
```
```{r}
plot.ts(test_BDO["Close"])
```

Using Predict() ung gawa nina deeks

```{r}
best = arima(BDO_TS_diff, order = c(4,0,0))
best
pred = predict(best, n.ahead= 30)$pred
pred
```

```{r}
actual_pred = ts(c(BDO_TS_diff, pred))
actual_pred = diffinv(actual_pred, differences = 1, xi = BDO_TS[1])
test_BDO_2 = ts(test_BDO["Close"])

plot(actual_pred, col = "blue", ylim = range(min(test_BDO_2, actual_pred), max(test_BDO_2, actual_pred)))
lines(test_BDO_2, col = "red")
legend("topleft", legend = c("Actual Prices", "Predicted Prices"), col = c("red", "blue"), lty = 1)

```



