---
title: "BDO Daily"
output: html_notebook
---

Loading Packages
```{r}
library("TSA")
library("tseries")
library("lmtest")
library("tidyverse")
library("forecast")
library("quantmod")
library("zoo")
```

Source of Data:l
Crude Oil prices from October 2001 to August 2020

Preprocessing the Data
```{r}
# Read the CSV data
Crude = read.csv("GoldUP.csv")
Crude = Crude[c(4:239),]
Crude_train = head(Crude, 228)
Crude_test = Crude["Crude_Oil"]
Crude
```
Create time series data
```{r}
Crude_ts = ts(Crude_train$Crude_Oil, start=c(2001, 1), end=c(2019, 12), frequency=12)
plot(Crude_ts)
```
```{r}
Crude_check=ts(Crude_test$Crude_Oil, start= c(2001,1), frequency=12)
plot(Crude_check)
```


Check for stationarity
```{r}
adf.test(Crude_ts)
```


Pre-process the price data. 
```{r}
Crude_ln_diff =diff(log(Crude_ts))
adf.test(Crude_ln_diff)
```
Since p<0.05, the price data is stationary. 

Check for autocorrelation
```{r}
Box.test(Crude_ln_diff, type = "Ljung")
```
Since p<0.05, we conclude that the price data is autocorrelated.

Modeling
```{r}
Acf(Crude_ln_diff, main = "ACF") # Check MA(1), MA(6)
Acf(Crude_ln_diff, type="partial", main = "PACF") # Check AR(1), AR(6)
```

We check the following models:

Lower AIC indicates the best fitting model.

```{r}
arima(Crude_ln_diff, order=c(1,0,0)) 
arima(Crude_ln_diff, order=c(6,0,0)) 
arima(Crude_ln_diff, order=c(0,0,1))
arima(Crude_ln_diff, order=c(0,0,6))
arima(Crude_ln_diff, order=c(1,0,1)) 
arima(Crude_ln_diff, order=c(6,0,1)) 
arima(Crude_ln_diff, order=c(1,0,6)) 
arima(Crude_ln_diff, order=c(6,0,6)) 
```

Define the models
```{r}
best = Arima(Crude_ln_diff, c(1,0,0))
ts.plot(Crude_ln_diff)
AR_fit = Crude_ln_diff - residuals(best)
lines(AR_fit, col = 2, lty = 2)
```

```{r}
predict = predict(best)
predict(best, n.ahead = 12)
```

```{r}
plot(Crude_ln_diff, xlim = c(2000, 2024))
AR_forecast = predict(best, n.ahead = 12)$pred
AR_forecast_se = predict(best, n.ahead = 12)$se
points(AR_forecast, type = "l", col = 'red')
points(AR_forecast - 2*AR_forecast_se, type = "l", col = 2, lty = 2)
points(AR_forecast + 2*AR_forecast_se, type = "l", col = 2, lty = 2)
```
```{r}
AR_forecast
final.ln.diff = ts(c(Crude_ln_diff,AR_forecast), start = c(2001, 1), frequency = 12)
final.ln = diffinv(final.ln.diff, differences = 1, xi = Crude_ln_diff[1])
final.ln
Crude_ln_diff
final = exp(final.ln)

plot(final, col = 2)
lines(Crude_ln_diff, col = 1, lty = 1)
```
```{r}
value = exp(cumsum(AR_forecast) + log(rev(Crude_ln_diff)[1]))
value_ts = ts(c(Crude_ts, value), start = c(2001, 1), end = c(2020, 12),frequency = 12)

plot(value_ts, col = 2)
lines(Crude_ts, col = 1)
```

```{r}
Crude_final_test = ts(Crude_test, start=c(2001, 1), frequency = 12)
plot(value_ts, col = "blue", ylim = range(min(Crude_final_test, value_ts), max(Crude_final_test, value_ts)))
lines(Crude_final_test, col = "red")
legend("topleft", legend = c("Actual Prices", "Predicted Prices"), col = c("red", "blue"), lty = 1)
```

Check the residuals. We want to see if they are autocorrelated.
```{r}
checkresiduals(best, test = "LB")
```
Since the p value > 0.05, we conclude that the residuals are not autocorrelated. (which is what we want)
Looking at the correlograms for the ACF and PACF, there are only a few lags which exceed the confidence threshold so the correlation between the residuals is not significant. The graph of the residuals shows that the residuals have a mean of around 0. 
