---
title: "psei_monthly data"
output: html_notebook
---

```{r}
library("TSA")
library("tseries")
library("lmtest")
library("tidyverse")
library("forecast")
library("quantmod")
library("ggplot2")
library("xts")
library("MLmetrics")
```

```{r}
psei = read.csv("PSEI.PS.csv")
psei$Date =  as.Date(psei$Date, format = "%d/%m/%Y") # format the dates as date objects
psei = psei[rowSums(is.na(psei)) != ncol(psei),] # remove empty rows
rownames(psei) = psei$Date
psei
```
```{r}
psei = psei["Close"]
psei_train = head(psei, 399)
psei_test = psei
psei = psei_train
psei_ts = ts(psei_train, start = c(1987, 2), frequency = 12)
#png(filename = "nonstationary.png")
plot(psei_ts, main = "Monthly Closing Prices for the PSEi")
#dev.off()
```
```{r}
psei
```


Check for Stationarity
```{r}
adf.test(psei_ts)
```
We can clearly see from the graph and from the AD test that the data is not stationary.

We perform log transformation and then difference the data.

```{r}
#png(filename="stationary.png")
psei_ts_ln = log(psei_ts)
psei_ts_ln_diff = diff(psei_ts_ln)
plot(psei_ts_ln_diff, main = "Stationary Time Series Plot after differencing and log transform")
#dev.off()
```

```{r}
adf.test(psei_ts_ln_diff)
```

The data is now stationary.

```{r}
Box.test(psei_ts_ln_diff, type = "Ljung")
```
The data is autocorrelated.

We can proceed with modeling.

```{r}
#png(filename="acf.png")
Acf(psei_ts_ln_diff, main = "ACF Correlogram") #MA(1)
#dev.off()
```


```{r}
#png(filename="pacf.png")
Acf(psei_ts_ln_diff, type = "partial", main = "PACF Correlogram") #AR(1)
#dev.off()
```

signficant lags at 1 for the acf/pacf.

```{r}
arima(psei_ts_ln_diff, order = c(1,0,0))
arima(psei_ts_ln_diff, order = c(0,0,1))
arima(psei_ts_ln_diff, order = c(1,0,1))
```
ARMA(1,0) has the lowest AIC.

Checking Residuals

```{r}
fit = arima(psei_ts_ln_diff, order = c(1,0,0))
checkresiduals(fit)
```

```{r}
residuals = fit$residuals
Acf(residuals)
Pacf(residuals)
```

At any lag, our residuals do not exhibit any autocorrelation. 


```{r}
psei_test_2 = ts(psei_test, start = c(1987, 2), frequency = 12)
model = Arima(psei_ts_ln_diff, order = c(1,0,0))
forecast_2yrs = predict(model, n.ahead = 24)$pred
final.ln.diff_2yrs = ts(c(psei_ts_ln_diff,forecast_2yrs), start = c(1987, 3), frequency = 12)
final.ln_2yrs = diffinv(final.ln.diff_2yrs, differences = 1, xi = psei_ts_ln[1])
final_2yrs = exp(final.ln_2yrs)

train = psei_ts
test = ts(psei_test_2[c(400:423)], start = c(2020,5), frequency = 12)
predicted = ts(final_2yrs[c(400:423)], start = c(2020,5), frequency = 12)

#png("forecast_2yrs.png")

plot(train, col = "blue", ylim = range(min(train, predicted), max(train, predicted)), main = "2-year forecast for the PSEi", xlab = "Year", ylab = "Price", xlim = c(1987, 2023), xaxt = "n")

lines(test, col = "#028A0F")
lines(predicted, col = "red")
legend("topleft", legend = c("Training Data", "Testing Data", "Forecasted Values"), col = c("blue", "#028A0F", "red"), lty = 1)
axis(1, at = seq(1987, 2026, by = 5), labels = seq(1987, 2026, by = 5))


# Close the PNG graphics device
#dev.off()
```


```{r}
#Finding the MAPE 
comparison = data.frame(
  actual = test,
  forecast = predicted
)
comparison
```
Computing the Mean Absolute Percentage Error
```{r}
mean(abs((comparison$actual-comparison$forecast)/comparison$actual)) * 100

```
MAPE of 6.62%. 

```{r}
psei_test = ts(psei_test, start = c(1987, 2), frequency = 12)
psei_test_ln = log(psei_test)
psei_test_ln_diff = diff(psei_test_ln)
model_2 = Arima(psei_test_ln_diff, order = c(1,0,0))
forecast_5yrs = predict(model_2, n.ahead = 60)$pred
final.ln.diff_5yrs = ts(c(psei_test_ln_diff,forecast_5yrs), start = c(1987, 3), frequency = 12)
final.ln_5yrs = diffinv(final.ln.diff_5yrs, differences = 1, xi = psei_ts_ln[1])
final_5yrs = exp(final.ln_5yrs)

train = psei_test
predicted = ts(final_5yrs[c(424:483)], start = c(2022,5), frequency = 12)

#png("forecast_5yrs.png")

plot(train, col = "blue", main = "5-year forecast for the PSEi", ylim = range(min(train, predicted), max(train, predicted)), xlab = "Year", ylab = "Price", xlim = c(1987, 2027), xaxt = "n")
lines(predicted, col = "red")
legend("topleft", legend = c("Training Data", "Forecasted Values"), col = c("blue", "red"), lty = 1)
axis(1, at = seq(1987, 2027, by = 5), labels = seq(1987, 2027, by = 5))


# Plot is saved as "forecast_plot.png"

# Close the PNG graphics device
#dev.off()
```


