---
title: "Jeremy Tan - Problem Set 3"
output: html_notebook
---



```{r}
library(TSA)
library(tseries)
library(forecast)
library(fGarch)
```


```{r}
data = dem2gbp
X = ts(data)
plot(X)
```
Assume that X_t is ARMA(0,0).

```{r}
Arima(X, order = c(0,0,0))
```

a. Verify that X_t is uncorrelated, but X_t^2 is serially correlated.

```{r}
Box.test(X, type = "Ljung")
```
Since p > 0.05, the X_t's are uncorrelated.

```{r}
Box.test(X^2, type = "Ljung")
```

Since p < 0.05, the X_t^2's are correlated.



Fit GARCH(1,1) and GARCH(1,2)

```{r}
garch11 = garchFit(~garch(1,1), data = X)
garch12 = garchFit(~garch(1,2), data = X)
```

```{r}
summary(garch11)
```

```{r}
summary(garch12)
```
For both models, mu is insignificant. Thus, we can let phi_0 be equal to zero. 

c. GARCH(1,1) has a lower BIC.

d. GARCH(1,2) has a higher log likelihood and a lower AIC, SIC, and HQIC.

e. I choose GARCH(1,2) as the best model.

Check the standardized residuals of ARMA(0,0) - GARCH(1,2)

```{r}
fit.vol = volatility(garch12)
fit.vol.sr = residuals(garch12) / volatility(garch12)
plot( fit.vol)
plot(fit.vol.sr)
```

```{r}
Box.test(fit.vol.sr, type = "Ljung")
```
Since p < 0.05, the standardized residuals are correlated. 
```{r}
acf(fit.vol.sr)
pacf(fit.vol.sr)
```


```{r}
Box.test(fit.vol.sr^2, type = "Ljung")

```

The squares of the standardized residuals are not correlated.