---
title: "R Notebook"
output: html_notebook
---

```{r}
library(TSA)
library(tseries)
library(forecast)
```

Item 2
```{r}
sales = read.csv("Item2.csv")
sales_ts = ts(sales)
```

Plot the data
```{r}
plot(sales_ts)
```
Do Log Transformation
```{r}
sales_ts_log = log(sales_ts)
plot(sales_ts_log)
```

Test for Stationarity
We do the adf test
```{r}
adf.test(sales_ts_log)
```

Since the p value is greater than 0.05, we fail to reject the null hypothesis. Thus, the time series is not stationary.

We do differencing.

```{r}
sales_ts_log_diff = diff(sales_ts_log)
```

```{r}
adf.test(sales_ts_log_diff)
```
Since the p value is greater than 0.05, we don't reject the null hypothesis. Thus, our time series is still not stationary.

We do differencing again.

```{r}
sales_ts_log_diff2 = diff(sales_ts_log_diff)
adf.test(sales_ts_log_diff2)
```

Since the p value is less than 0.05, we reject the null hypothesis. The twice differenced time series is now stationary.

```{r}
plot(sales_ts_log_diff2)
```
Check for autocorrelation
```{r}
Box.test(sales_ts_log_diff2, type = "Ljung")
```
Since p < 0.05, the time series is autocorrelated.

Fit an ARIMA(1,2,1) model to sales_ts_log

```{r}
arima_121 = Arima(sales_ts_log, order = c(1,2,1))
arima_121
```
The prediction interval of phi_1 includes 0 so the phi_1 term can be dropped.

Let's test out the ARIMA(0,2,1) model.

```{r}
arima_021 = Arima(sales_ts_log, order = c(0,2,1))
arima_021
```

ARIMA(0,2,1) has lower AIC and BIC compared to ARIMA(1,2,1) so we use ARIMA(0,2,1) for forecasting.

```{r}
sales.f = forecast(arima_021, h = 3)
sales.f
```
```{r}
sales.f.actual = exp(sales.f$mean)
sales.f.actual
```


Item 3.

```{r}
rate = read.csv("Item3.csv")
rate_ts = ts(rate)
```

```{r}
plot(rate_ts)
```
We can notice there is a seasonal aspect and a trend.

Do log transform and differencing.

```{r}
rate_ts_log = log(rate_ts)
rate_ts_log_diff = diff(rate_ts_log)
```

```{r}
plot(rate_ts_log_diff)
```
```{r}
acf(rate_ts_log_diff, lag = 100)
```
```{r}
pacf(rate_ts_log_diff, lag = 100)

```
For the seasonal component, we have a sARMA(1,0)_12 model.

We test out two different models:
1. sarima(1,1,2) x (1,1,0)_12
2. sarima(2,1,1) x (1,1,0)_12

```{r}
sarima.arma12 = Arima(rate_ts_log, order = c(1,1,2), seasonal = list(order = c(1,1,0), period = 12))
sarima.arma12
```

```{r}
sarima.arma21 = Arima(rate_ts_log, order = c(2,1,1), seasonal = list(order = c(1,1,0), period = 12))
sarima.arma21
```
```{r}
rate.f = forecast(sarima.arma21, h = 3)
rate.f
```
```{r}
exp(rate.f$mean)
```

Item 4
```{r}
A = matrix(c(5/32, 3/32, 9/160, 27/800, 3/32, 5/32, 3/32, 9/160, 9/160, 3/32, 5/32, 3/32, 27/800, 9/160, 3/32, 5/32), ncol = 4, byrow = T)
A
solve(A)
```

```{r}
B = matrix(c(3/32, 9/160, 27/800, 81/4000), ncol = 4)
Y = matrix(c(-0.1, 0, 0, 0.6), ncol = 1)
B %*% solve(A) %*% Y
```
```{r}
C = matrix(c(9/160, 27/800, 81/4000, 243/20000), ncol = 4)
C %*% solve(A) %*% Y
```

```{r}
B %*% solve(A) %*% t(B)
```
```{r}
C %*% solve(A) %*% t(C)
```

