---
title: "R Notebook"
output: html_notebook
---



```{r}
library(TSA)
library(tseries)
data("usd.hkd")
forex = usd.hkd$hkrate
library(forecast)
arma21 = Arima(forex, order = c(2,0,1))
arma21
```

```{r}
forecast(arma21, h = 5)
```
Gives the point forecast and confidence intervals

```{r}
plot(forecast(arma21, h = 5))
```

```{r}
tsdiag(arma21)
```

The residuals should be uncorrelated.
The ACF should be zero for h > 0.
In this example, the residuals are not correlated.

```{r}
acf(arma21$residuals)
```
```{r}
jarque.bera.test(residuals(arma21))
```
Since p < 0.05, we reject the null hypothesis that the residuals come from a normal distribution. We conclude that the residuals are white noise.
```{r}
library(itsmr)
```
```{r}
strikes
```
```{r}
plot(strikes)
```

Since the values are large, we log transform.

```{r}
strikes.log = log(strikes)
detach("package:itsmr", unload = TRUE)
```
We unload itsmr since it interferes with the forecast function.

```{r}
library(forecast)
plot(strikes.log)
```

```{r}
adf.test(strikes.log)
```
Since p > 0.05, we do not reject the null hypothesis. The time series is not stationary.

```{r}
strikes.log_diff = diff(strikes.log)
adf.test(strikes.log_diff)
```

The p value is still greater than 0.05 so it is not stationary.

```{r}
strikes.log_diff2 = diff(strikes.log_diff)
adf.test(strikes.log_diff2)
```
Since p < 0.05, the time series is now stationary.

```{r}
Box.test(strikes.log_diff2, type = "Ljung")
```
Since p < 0.05, we reject the null hypothesis. The time series is autocorrelated.

```{r}
acf(strikes.log_diff2) #MA(1)
```
```{r}
pacf(strikes.log_diff2) # AR(2)
```

```{r}
ma1 = Arima(strikes.log_diff2, order = c(0,0,1))
ar2 = Arima(strikes.log_diff2, order = c(2,0,0))
arma21 = Arima(strikes.log_diff2, order = c(2,0,1))

```

```{r}
ma1
ar2
arma21
```

Snce AR(2) has the lowest AIC, BIC, and AICc, we select AR(2) as the best fitting model.

```{r}
ar2.d = Arima(strikes.log, order = c(2,2,0))
```
The order has middle index, 2 because we differenced twice.

```{r}
ar2.d
```
```{r}
ar2 = Arima(strikes.log_diff2, order = c(2,0,0), include.mean = F) # we indicate mean zero
ar2
```
For phi_2, we have the interval -0.78 +- 1.96(0.1368) or (-1.048, -0.51187). The interval does not include zero so the coefficient is significant.

```{r}
plot(forecast(ar2, h = 5)) 
```
Forecast of the original data

```{r}
plot(forecast(ar2.d, h = 5))
```

```{r}
tsdiag(ar2.d)
```
For all h > 0, the ACF is zero so the residuals are uncorrelated.

Seasonal Models

```{r}
air = AirPassengers
air
```
```{r}
plot(air)
```
The data has seasonality and follows a trend.

```{r}
adf.test(air)
```

The adf test doesn't work with seasonal data.

```{r}
air.l = log(air)
air.l.d = diff(air.l)
```

```{r}
plot(air.l.d)
```

```{r}
plot(decompose(air.l))
```
```{r}
plot(decompose(air.l.d))
```
```{r}
Box.test(air.l.d, type = "Ljung")
```
The time series is autocorrelated.

```{r}
acf(air.l.d)
```
```{r}
acf(air.l.d, lag = 100)
```

Every twelve months, the ACF is significant.

```{r}
pacf(air.l.d, lag = 100)
```
The PACF cuts off at 12.
The ACF cuts off at 12, 24, 36, ...

We have an ARMA(1,0)_12 model.

We do seasonal differencing.

```{r}
air.l.d.d12 = diff(air.l.d, 12)
plot(air.l.d.d12)
```
```{r}
plot(decompose(air.l.d.d12))

```

We have sARIMA( ,1, ) x (1,1,0)_12

We find the non-seasonal components.

We test out AR(1), MA(1), ARMA(1,1)

```{r}
sarima.ar1 = Arima(air.l, order = c(1,1,0), seasonal = list(order = c(1,1,0), period= 12))
sarima.ar1
```

```{r}
sarima.arma1 = Arima(air.l, order = c(1,1,1), seasonal = list(order = c(1,1,0), period= 12))
sarima.arma1
```
```{r}
sarima.ma1 = Arima(air.l, order = c(0,1,1), seasonal = list(order = c(1,1,0), period= 12))
sarima.ma1
```
We select MA(1) as it has the lowest criterion.

Final model is sARIMA(0,1,1) X (1,1,0)_12.

```{r}
plot(forecast(sarima.ma1, h = 12))
```

```{r}
air.f = forecast(sarima.ma1, h = 12)
air.f
```

```{r}
air.f$mean
```
```{r}
preds = exp(air.f$mean)
preds

```


```{r}
air_preds = ts(c(air, preds), start = c(1949, 1), frequency = 12)
```

```{r}
plot(air_preds)
```

```{r}
plot(forecast(sarima.ma1, h = 12))
```

